---
title: "Leverging PubMed records for business applications"
author: "Damiano Fantini"
date: "May 28, 2018"
output:
  pdf_document:
    latex_engine: xelatex
  html_document:
    df_print: paged
vignette: >
  %\VignetteIndexEntry{Leverging PubMed records for business applications}
  %\VignetteEngine{knitr::rmarkdown}
  %\usepackage[utf8]{inputenc}
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(easyPubMed)
library(businessPubMed)
```

## Installing the businessPubMed package

The `businessPubMed` package can be installed from GitHub (via devtools). It requires the latest version of easyPubMed (version 2.5 available on CRAN, or version 2.7 available on GitHub). The `devtools` package is available on CRAN.


```{r install01, include=T, eval=FALSE}
# install the package via > devtools::install_github("dami82/businessPubMed")
library(devtools)
install_github("dami82/easyPubMed")
install_github("dami82/businessPubMed")

library(easyPubMed)
library(businessPubMed)
```

## Getting started

Here, we showcase how to use `businessPubMed` to retrieve email addresses from PubMed records. Specifically, we will: i) retrieve recent records about a subject of interest; ii) extract email addresses of contributing and corresponding authors from these records; iii) automatically prepare a message with a topic- and publication-specific mock offer (real names/adds are masked in the vignette) for a targeting campaign. In this example, we are querying PubMed for works published in the United States between 2012 and 2017 on the topic: "T cells".

```{r run01, include=T, eval=T}
# define a research subject
res.subject <- "t cell"

# define a PubMed Query string
my.query <- paste('"', res.subject, '"[TI]', sep = '') 
my.query <- paste(my.query, 'AND (USA[Affiliation] OR "United States")' , sep = '')
my.query <- paste(my.query, 'AND ("2013"[EDAT]:"2017"[EDAT])')

# Show query string
cat(my.query)
# define the regex filter
data("countries")
countries <- countries[countries != "United States of America"]
countries.filter <- gsub("[[:punct:]]", "[[:punct:]]", toupper(countries))
countries.filter <- gsub(" ", "[[:space:]]", countries.filter)
countries.filter <- paste("(",countries.filter,")", sep = "", collapse = "|")
countries.filter <- paste("(UK)|", countries.filter, sep = "")

```
```{r run02, include=T, eval=F}
# Show filter string
cat(countries.filter)
```
```{r run02real, include=T, eval=T, echo=FALSE}
# Show filter string
cat(paste(substr(countries.filter, 1, 85), "...", sep = ""))
```

```{r run03, include=T, eval=T}
# Check time t_init
t_init <- Sys.time()

# extract data from PubMed
my.data <- businessPubMed::extract_pubMed_data(pubMed_query = my.query, 
                                               batch_size = 1000,
                                               affi_regex_exclude = countries.filter)

# Check time t_final
t_final <- Sys.time()

# How many unique documents?
my.data$params$pubMed_id_list$Count
# How many unique addresses?
sum(!is.na(unique(my.data$data$address)))
# How many unique emails?
sum(!is.na(unique(my.data$data$email)))
# How long did this analysis take?
print(t_final - t_init)
# Automated message composition - filter missing email addresses
contact.DF <- my.data$data[!is.na(my.data$data$email),]

# Automated message composition - process one record
# You may want to wrap this in a `for (i in 1:nrow(contact.DF)) {...}` loop
i = 1
tmp.addressee <- contact.DF[i,]
```
```{r run04, include=F, eval=T, echo=F, include=F}
tmp.addressee$firstname <- "John"
tmp.addressee$lastname <- "Doe"
tmp.addressee$title <- "The identification of key markers of lymphoproliferative disorders"
tmp.addressee$journal <- "American Journal of Cancer and Immunity"
tmp.addressee$address <- "Department of Internal Medicine, State University, Springfield, USA"
tmp.addressee$email <- "john.doe@sprinfieldstateuniv.edu"
```
```{r run05, include=T, eval=T}
# Compose the message for the targeting campaign
mail.text <- paste("\n", tmp.addressee$address, "\n\n", "Dear Dr. ", 
                   tmp.addressee$lastname, " ", tmp.addressee$firstname, 
                   ", \n\n    I really appreciated reading your ", 
                   tmp.addressee$year ," manuscript that was published\non ", 
                   tmp.addressee$journal, ". I believe that your work on\n", 
                   gsub("[[:punct:]]$", "", tolower(tmp.addressee$title)), 
                   "\nmay really benefit from the products included in ", 
                   "our offer. \nFor example, we sell a product specifically", 
                   " designed for research on ", res.subject, "\nand I am ", 
                   "happy to include a special offer that you may find ",
                   "interesting. \n[...] \nFeel free to contact me at the", 
                   " following address XXXXX... \n \nRegards. \n \n", 
                   "Frank Doe \nCompany XYZ", sep = "")

# Here you should add code for preparing and sending an email.
# However, this goes beyond the scope of this vignette.
# Here, we display the text of our mock email
cat(mail.text)
```

The whole analysis can be performed reasonably fast (about 4 mins per 1000 records, using a standard laptop). Moreover, parallelization may be exploited to further speed up the processing time. 
Other than for targeting campaigns, `businesPubMed` can be used for a wide panel of similar application. I hope you found this vignette useful.
Thanks for using `easyPubMed` and `businessPubMed`.

Damiano Fantini (Copyright 2018)

## Session Info

```{r run06}
print(sessionInfo())
```

